\begin{workpackage}[id=WP_active,wphases=0-48,
  short=Active Particle Suspensions,% for Figure 5.
  title=Probing Active Particle Suspensions with Colloids and Polymers,
  lead=Leipzig,
  LeipzigRM=12]

\begin{wpobjectives}
The objectives of this WP are:
  \begin{compactitem}
      \item \textbf{Non-Equilibrium Equations of State (NEOS):} explore the existence of macroscopic non-equilibrium equations of state for active Brownian particle suspensions with microscopic many-particle models
 \item \textbf{Active Crowding:} elucidate by analytical model calculations how mechanical and chemical activity (such as by molecular motors in the cytoplasm of animal cells) affects the statistical mechanics of embedded biopolymers (such as protein fibers or DNA)
      \item \textbf{Noise Temperature:} validate the concept of a frequency-dependent noise temperature in non-equilibrium fluctuation-dissipation relations and the failure of Onsager's regression hypothesis for colloidal probes in non-isothermal solvents by experiments and computer simulations
      \item \textbf{Active-Particle Suspensions:} scrutinize tentative non-equilibrium equations of state by a combination of experiments and all-atom non-equilibrium molecular-dynamics simulations
  \end{compactitem}
\end{wpobjectives}

\begin{wpdescription}
  In equilibrium, information about the macroscopic state of a medium can be deduced via rather simple measurement procedures using simple probes and then be summarized in an equation of state. 
  In the future, something similar may become possible even far from equilibrium, and revolutionize our technological abilities, but this will require much more sophisticated probing and analyzing procedures that are not yet generally understood. 
  Far from equilibrium, fluctuations are non-universal and standard thermodynamic notions, such as temperature and pressure, become ambiguous. 
  This has many unintuitive practical consequences, which defy conventional thermodynamic wisdom and potentially invalidate and doom to failure well-established standard technological procedures. 

  For example, a colloidal particle in a non-isothermal solvent cannot be used for Brownian thermometry in the conventional way, and also the proper theoretical treatment of its systematic, thermophoretic ``swimming'' is the subject of ongoing debates. 
  Also, the mechanical pressure will generally hinge on the range and symmetry of the particle-probe interactions: if an active-particle suspension is compressed by a piston, the amount of work needed to reach a prescribed amount of compression depends on the type of piston used (e.g.\ how hard or soft its surface is), and the pressure can be macroscopically anisotropic and heterogeneous in absence of macroscopic stationary fluxes. 
  
Similarly, any biopolymer in a living cell actually represents a complex (or smart) probe that scans its non-equilibrium mesoscopic environment with its many internal degrees of freedom. 
  The myriads of motile proteins and motors in a living cell cause deviations of the function of its constituents with respect to conventional equilibrium predictions. 
  We call this effect ``active crowding'', in analogy with the much studied ``molecular crowding''. 
  We could recently quantitatively demonstrate by theory and simulations that the effect of a quenched static crowding onto an embedded semiflexible polymer can effectively be captured by a relatively simple parameter renormalization. 
  Here we ask: can something similar be achieved for active crowding? 

  There is currently no consensus on how the fundamental obstacles to the formulation of non-equilibrium equations of state can systematically be overcome, as underscored by a number of research articles published this year, which make claims towards either the existence or non-existence of NEOS for active-particle suspensions.
  These complications severely limit a first-principles approach to living systems and technological applications. 
 Serious conceptual problems need to be solved before, say, active particle suspension will become the basis for new materials with built-in actuation mechanisms, for fluid mixing, viscosity reduction, or shifting the glass transition in industrial processing, or for noninvasive surgeries and targeted drug delivery. 
  We address them as systematically as we can (though not comprehensively) with the above objectives 1 \& 2 (theoretically) and 3 \& 4 (experimentally and by massively parallel computer simulations). 
 
\end{wpdescription}

\begin{tasklist}

  \begin{task}[title=Non-Equilibrium Equations of State (NEOS),id=task1,PM=15,lead=Leipzig,wphases=0-30!0.5]
    As a paradigm for a non-equilibrium medium we consider the (experimentally vast) class of active-particle suspensions. 
    Currently studied microscopic models all have special explicit, and often additional implicit symmetries and moreover violate fundamental physical principles, such as momentum conservation. 
    They are therefore sometimes characterized as ``dry active matter'', as opposed to (usually) more realistic ``wet active matter''.
    In our approach, we follow this common procedure (as e.g.\ in actual papers by Solon et al.) with the aim to successively add elements of ``wet reality'' later on. 
    Our starting point is a many-body Langevin equation with a Markovian Gaussian noise (describing the coupling to an equilibrium heat bath) and individual propulsion forces for the particles representing, on an all-atom-level, the active fluid far from equilibrium. 
    We then let the fluid interact perturbatively with one large colloidal probe particle to deduce its effective friction kernel, and with two or more probe particles in order to detect their induced mutual interactions, non-gradient forces, and a possible breaking of the action-reaction principle on the coarse-grained probe level.
    Non-equilibrium response theory lets us, in particular, include frenetic contributions in the response functions. 
    Of particular interest to us is the mechanical pressure exerted by an active fluid on a probe, in particular if we relax the very special geometric symmetries tacitly imposed in previous approaches. 
    To add elements of ``wet reality'', we will further account for more realistic particle-probe interactions, as appropriate e.g.\ for hot Janus particles. 
    The required ingredients are currently being developed by us, using hydrodynamic theory, simulations, and experiments, in an ongoing joint project within the German priority program ``Microswimmers''.     
     
It is to date still not clear, whether systems in which the macroscopic non-equilibrium is caused by self-propulsion of the fluid particles are fundamentally different from those exposed to an external driving, e.g.\ by shearing or laser heating. 
    We plan to elucidate this question in collaboration with our theoretical colleagues from the node in Stuttgart by establishing a connection of our particle-based approach with their field theoretical approach.

  \end{task}

  \begin{task}[title=Active Crowding,id=task2,PM=15,lead=Leipzig,wphases=12-42!0.5]
    A first attempt to estimate how the activity in a living cell affects the physical behavior and function of its constituents was put forward by Mikhailov and Kapral in a paper in PNAS, earlier this year. 
    They predict that activity will, via hydrodynamic interactions, enhance the diffusion of passive globular probes or cell organelles.  
    Yet, it remains open, whether such perturbative results get enhanced (as in equilibrium) or largely diminished on a non-perturbative level, if one considers, for example, that the active stirring may affect the effective friction of a probe particle, or that motors or other sources of non-equilibrium do often act directly on cytoskeletal polymers or their weak transient networks.    
    Armed with our previous experience with the derivation of effective friction coefficients under far-from-equilibrium conditions, we therefore aim to find out how the activity affects the effective non-equilibrium friction of a probe, and whether it enhances or counteracts the perturbatively predicted diffusivity. 
    Further, we want to work out, from first principles, how three different types of active crowding affect the fluctuation spectrum and dynamics of a semiflexible test polymer: a laser-heated metal nanoparticle tracer attached to a micron-sized semiflexible polymer such as an actin microfilament or a microtubulus; power strokes exerted by attached molecular motors and inducing backbone tension spikes along the polymer contour; ``inelastic'' unbinding of weak bonds along the test polymer, which are the ubiquitous cement of cytoskeletal structures, under static or dynamic loads. 
    For these tasks we can rely on our strong track record of first-principle analytical studies of the non-linear non-equilibrium dynamics of biopolymers, non-quasistatic bond kinetics under strong external forcing, and hot Brownian motion. 
    Think for example how the free recoil of a stretched DNA molecule, which is solely driven by fluctuations (and for which we have recently demonstrated a perfect parameter-free match of theory and experiment under isothermal conditions) is affected by the various types of active excitations.  
    Finally, it will be of major interest to clarify to what extent our models of active crowding admit a formal mapping to the situation in non-isothermal simple solvents, via frequency-dependent noise temperatures and effective friction coefficients.
\end{task}


 \begin{task}[title=Noise Temperature,id=task2,PM=15,lead=Leipzig,wphases=12-42!0.5]    
   We have recently established a tractable ``drosophila'' for the central theme of the proposal, i.e.\ probing macroscopic non-equilibria, in terms so-called non-isothermal (or ``hot'') Brownian motion. 
   Using analytical calculations, we have demonstrated that the phenomenon is amenable to an analytical framework that extends conventional equilibrium statistical mechanics formalisms by introducing a generalized notion of temperature, chiefly a frequency-dependent temperature for the Brownian noise strength (``noise temperature''). 
   Corner stones of statistical mechanics, such as the fluctuation-dissipation theorem, can thereby be generalized to conditions far from equilibrium.
   However, these results, which are relevant to metal-nanoparticle tracking and trapping technologies and emerging nanoscopic Brownian heat engines, could so far not be validated by experiments and microscopic computer simulations, beyond the easily accessible Markov limit. 
   We therefore propose to develop a high-speed experimental measurement setup capable of detecting inertial effects in the Brownian motion of hot metal nanoparticles and thereby to detect the predicted increase of the apparent Brownian temperature of their short-time dynamics. 
   This requires to follow the fluctuations of a heated Brownian particle on a nanosecond timescale with a positional accuracy of picometers and below to measure the (modified) Maxwell velocity distribution, a task that has only very recently been achieved under equilibrium conditions, using latest optical-tweezers technology. 
   Extending the method to conditions far from equilibrium requires substantial additional technological developments, e.g.\ due to heating effects and radiation pressure. 
   A dedicated optical trap with dual beam heating and a detection bandwidth of about 50 MHz will therefore be constructed. 
      
Further, to test the theory also by numerical simulations, we will quantify the deviations from the idealized theory developed for incompressible fluids to Lennard--Jones solvents, as used in our efficient massively parallel computer simulations, which exhibit a substantial compressibility. 
   A successful comparison of the results from non-equilibrium molecular-dynamics simulations and theory should also provide a template for analyzing future experiments realizing extreme probe temperatures. 
\end{task}

  \begin{task}[title=Active-Particle Suspensions,id=task2,PM=15,lead=Leipzig,wphases=12-42!0.5]
    Technological applications of active-particle suspensions and systematic tests of uncontrolled approximations that are currently unavoidable in all theoretical studies are currently severely hampered by a lack of control over these systems. 
    Therefore, dedicated experimental investigations are still very rare and some of their results might be specific to the experimental realization of the self-propulsion mechanism.
    We have recently established innovative force-free control systems, so-called photon nudging and thermophoretic trapping technologies, which deliberately exploit the non-equilibrium transport caused by non-isothermal conditions. 
    Thereby we can e.g.\ construct soft walls of passive or active particles that act as semi-permeable membranes.
    In combination with conventional manipulation tools such as structured steric confinements, light forces, electromagnetic fields, and gravitation, these enable us to realize more ambitious and better controlled setups involving self-propelled particles than hitherto possible. 
    We will predominantly use hot Janus particles in water controlled by laser heating, which allows for a continuous tuning of the strength of the activity in space and time. 
    While recent experimental work with a highly symmetric active-particle suspension setup has suggested that activity can be accounted for in the thermodynamic bulk-pressure equation of state by a renormalized temperature and renormalized interactions, we want to more specifically target possible failures of this scheme. 
    Using a variety of designs of potential confinement together with our force-free control technology, we will elucidate to what extent wall interactions (rather than the intrinsic asymmetry of the propulsion mechanism) are the fundamental source of time-reversal asymmetry at the coarse-grained level, and thus relevant for the breaking of detailed balance and ensuing rectification (``ratchet'') effects. 
    Thereby, and by accompanying numerical simulations, we aim to detect discrepancies between the thermodynamic bulk pressure and the actual mechanical wall pressure. 
\end{task}


\end{tasklist}

\begin{wpdelivs}
  \begin{wpdeliv}[due=12,id=mydeliv1,dissem=PU,nature=DEM,lead=Leipzig]
      {First deliverable, after 1 year.}
  \end{wpdeliv}
  \begin{wpdeliv}[due=24,id=mydeliv2,dissem=PU,nature=DEM,lead=Leipzig]
      {Second deliverable, after 2 years.}
\end{wpdeliv}
\end{wpdelivs}

\end{workpackage}
