\begin{sitedescription}{ULEI} \label{desc:ULEI}

 Universität Leipzig was founded in 1409 and is the second oldest university in Germany where teaching has continued without interruption. 
  Today it offers a wide spectrum of academic disciplines at 14 faculties with more than 150 institutes. 
  It is a member of the German U15, a strategic alliance of 15 major German research universities. 
  The node participants are located at the Institute for Experimental Physics I and at the Institute for Theoretical Physics, respectively, which are parts of the Faculty of Physics and Earth Sciences. 
  The faculty also comprises institutes for meteorology, geology, geophysics and geography. 
  It counts among the leading faculties in terms of research output and external funding, within the university, and hosts several ERC grantees. 
  It is also the first faculty of the university that recently got its research activities evaluated by an external board of international experts. 
  The physics institutes make major contributions to several collaborative research centers (SFBs) funded by the German Science Foundation (DFG) and to the interdisciplinary Graduate School ``Leipzig School of Natural Sciences -- Building with Molecules and Nano-objects'' (www.buildmona.de), founded by a grant from the German Excellence Initiative, which has so far enrolled close to 200 PhD candidates. 
  They maintain collaborations and joint grants with a large number of independent international and local research institutes for fundamental and applied science and industrial partners. 

\subsubsection*{Curriculum vitae of the investigators}

\input{CVs/Klaus.Kroy.tex}
\input{CVs/Frank.Cichos.tex}

\begin{participant}[type=res,PM=48,salary=5500]{NN}
A postdoc will be hired to work on the project. We aim to hire someone who has a strong
background in theoretical statistical mechanics and a diverse research experience. The person will keep contact to the other project nodes and the experimental project partners.
\end{participant}

\begin{participant}[type=res,PM=36,salary=5500]{NN}
A postdoc researcher is required to carry out the experiments on hot Brownian motion at nanosecond timescales. The postdoc should have a solid background in optical tweezer experiments and fast positional detection. He shall work closely with the theoretical project partners and keep close contact to the other nodes.
\end{participant}

\subsubsection*{Publications, achievements}

\begin{compactenum}
\item M. Gralka, K. Kroy, Inelastic mechanics: A unifying principle in biomechanics. 
Biochimica et Biophysica Acta (BBA)--Molecular Cell Research 2015

\item S. Schöbl, S. Sturm, W. Janke, K. Kroy, Persistence-Length Renormalization of Polymers in a Crowded Environment of Hard Disks.
Physical Review Letters {\bf 113} 238302 (2014).

\item J. T. Bullerjahn, S. Sturm, K. Kroy, Theory of rapid force spectroscopy. Nature Communications {\bf 5} 4463 (2014).

\item O. Otto, S. Sturm, N. Laohakunakorn, U. F. Keyser, K. Kroy, Rapid internal contraction boosts DNA friction.
Nature Communications {\bf 4} 1780 (2013).

\item D. Chakraborty, M. V. Gnann, D. Rings, J. Glaser, F. Otto, F. Cichos, K. Kroy, 
Generalised Einstein relation for hot Brownian motion. EPL (Europhysics Letters) {\bf 96} 60009 (2011).

\item D. Rings, R. Schachoff, M. Selmke, F. Cichos, K. Kroy. 
Hot Brownian Motion.  Physical Review Letters {\bf 105},  090604 (2010). According to Google Scholar, this article has been cited 81 times. It highlights the interest in fundamental non-equilibrium fluctuation dissipation theorems and its experimental verification.

\item M. Braun, A. Bregulla, K. Günther, N. Mertig, F. Cichos Single Molecules Trapped by Dynamic Inhomogeneous Temperature Fields Nano Lett. 15 5499 (2015).

\item A. Bregulla, H. Yang, F. Cichos Stochastic Localization of Micro-Swimmers by Photon Nudging ACS Nano 8 6542 (2014).

\item M. Selmke, M. Braun, F. Cichos Photothermal Single Particle Microscopy: Detection of a Nanolens ACS Nano 6 2714 (2012).

\end{compactenum}

\subsubsection*{Previous projects or activities}

\begin{compactenum}
\item Project lead (speaker) of the research unit 877 ``From Local Constraints to Macroscopic Transport'' of the German Science Foundation (DFG).
%
\item Organization of the "Hot Nanostructures" workshop series.
This workshop has been initiated in 2011 by the two principle investigators in Leipzig and is continued every two years since then. It is focused especially on problems of non-equilibrium physics highlighting theory and experiments at inhomogeneous temperatures. The workshop will return in 2017 to Leipzig.

\item Partner in international collaborative projects (Agence National de la Recherche, ANR - German Science Foundation, DFG)
\end{compactenum}

\subsubsection*{Significant infrastructure}

\end{sitedescription}

\begin{draft}
\vspace{1cm}\TOWRITE{PAR1P1}{Complete check list below -- delete completed items if you wish}

\begin{verbatim}
- [ ] checked that sum of person months put into finance request is
  the same as sum of person months associated with the Work Packages
  (in proposal.tex, as defined as part of the \begin{workpackage}"
  command.
  
- [ ] completed site specific resource summary in resources.tex,
  including table of non-staff costs.

\end{verbatim}
\end{draft}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End: 
