\begin{sitedescription}{USTUTT} \label{desc:USTUTT}

{\bf Universit채t Stuttgart -- A Research University of International Standing:}\\
The Universit채t Stuttgart lies right in the centre of the largest high-tech region of Europe. We are surrounded by a number of renowned research facilities and have such global players as Daimler or IBM as our neighbours. We were founded in 1829 and over the years this technical institution has developed to the research intensive university that it is today. Our main emphasis is on engineering and the natural sciences.
%
Indicators of our excellent status are the two projects that were successful in the recent {\it Excellence Initiative} sponsored by both the Federal and the State governments. One project is the Cluster of Excellence {\it Simulation Technology} and the other, the Graduate School {\it Advanced Manufacturing Engineering}.

{\bf Experience with EU research funding:}\\
The University of Stuttgart has extensive experience with the various funding programs of the European Commission and has been the {\it leading German university in FP6}, both in number of projects (184) and in terms of funding (54 Mio. \euro). In FP7, it was yet again {\it among the most successful German universities} with 246 projects funded and a total budget of 94 Mio. \euro.

\subsubsection*{Curriculum vitae of the investigators}

\input{CVs/Matthias.Kruger.tex}
\input{CVs/Clemens.Bechinger.tex}

\begin{participant}[type=res,PM=48,salary=5500]{NN}
We will hire a Postdoc with a strong theoretical background in (non-equilibrium) statistical
physics.
\end{participant}
\begin{participant}[type=res,PM=36,salary=5500]{NN}
We will hire a Postdoc with experimental experience concerning manipulation and measurement
of colloidal systems.
\end{participant}

\subsubsection*{Publications, achievements}

\begin{compactenum}
\item Matthias Kr체ger developped theoretical frameworks predicting properties of
far-from-equilibrium fluids, especially near interfaces and for glassy viscoelastic
materials and he contributed to the description of electromagnetic fluctuations in
non-equilibrium situations (e.g. in Physical Review Letters).
\item Clemens Bechinger recently published experimental articles on (non-equilibrium)
stochastic thermodynamics, self propelled swimmers, critical Casimir forces, quasicrystals,
a micrometer sized heat engine and nanotribology (in Nature, Nature Materials, Nature
Physics, Nature Comm. and Physical Review Letters).
\end{compactenum}

\subsubsection*{Previous projects or activities}

\begin{compactenum}
\item During the past five years, Clemens Bechinger was funded by several programs, e.g. by
two {\it Individual Grants} and a {\it Priority Progam} of the German Research Foundation
(DFG). His research was also supported as a {\it Max-Planck-Fellow} as well as by the {\it
  Initial Training Networks} of the Marie Curie Actions.
\item Matthias Kr체ger held an {\it Individual Fellowship} of the DFG starting 2010 and is
currently supported through a junior group grant ({\it Emmy Noether Program}) funded by the
DFG.
\end{compactenum}

\subsubsection*{Significant infrastructure}
The research group in Stuttgart is equipped for the study of active particles systems, of
visco-elastic media and of colloids in general.
%
This infrastructure plays a critical role in the development of \TheProject.

\end{sitedescription}
