\begin{sitedescription}{TUE}

TUE is an internationally leading research university specialised in engineering science and technology. TUE is a constant presence in the top of various rankings, all acknowledging in particular the research carried out in collaboration with the industry, and currently manages more than 150 EU-funded projects. TUE offers 50 different educational programs to about 9000 students of various levels, and employs more than 3000 academic and administrative personnel.

The TUE research team in this proposal is part of both the Department of Mathematics and Computer Science and the Institute for Complex Molecular Systems. 
The Department Mathematics and Computer Science (MCS) unites all activities on campus in the classical scientific disciplines of mathematics and computer science. Both research and education have recently been evaluated as excellent by international committees. 40\% of the employees at MCS are of non-Dutch origin, thus creating a truly multi-cultural environment. 

The Institute for Complex Molecular Systems (ICMS) is a hotbed for interdisciplinary interaction in research and education across the university. Its mission is to become the leading international multidisciplinary institute for research and education in the area of the engineering of complex molecular systems.

The joint affiliation with both the Department of Mathematics and Computer Sciences and the ICMS provides the TUE research team with both excellent mathematical infrastructure and broad experimental embedding.


\subsubsection*{Curriculum vitae}

% Curriculum of the personnel at this institution. This includes
% to-be-hired people for which there is a tentative candidate.

\input{CVs/Mark.Peletier.tex}

\input{CVs/Adrian.Muntean.tex}
% For other to-be-hired person, please include here something like:
% \begin{participant}[type=res,PM=3,salary=5900]{NN}
%  <a _short_ description of the qualifications of whom you want to hire>
% \end{participant}

\begin{participant}[type=res,PM=48]
A PhD student will be hired for the duration of the project that matches the duration of a
PhD thesis in the Netherlands. 
%
The opening will be international and open to competent candidates from any
origin.
\end{participant}


\subsubsection*{Publications, products, achievements}

\begin{compactenum}
\item {In a series of papers, starting with \emph{Adams et al., Communications in Mathematical Physics, 307:791 (2011)}, Peletier and co-authors identified and explored the deep relations between gradient flows on one hand and large deviations of stochastic processes on the other. These relations create new understanding of the mathematical structures describing physical phenomena at the mesoscale, and point the way towards understanding the strongly nonequilbrium systems of this proposal.}
\item {M. A. Peletier, G. Savar\'e, and M. Veneroni, From diffusion to reaction via $\Gamma$-convergence, SIAM Journal on Mathematical Analysis, 42(4), pp. 1805--1825, 2010.
This paper has sparked renewed interest in the analysis community in the old problem of a stochastic particle escaping from a potential well, by introducing a new method to combine reactive and diffusive effects. It was selected as a SIGEST paper in SIAM Review 54(2) in 2012.}
\item {Peletier is a co-founder and board member of the ICMS, and both Peletier and Muntean are prominent members of this interdisciplinary institute.}
\end{compactenum}

\subsubsection*{Previous projects or activities}

\begin{compactenum}
\item {Peletier is and has been PI on many national and international projects, including an EU Seventh-framework ITN project `Fronts and Interfaces in Science and Technology'.}
\item {Peletier and Muntean have together and separately organized over 50 meetings, workshops, and conferences.}
\end{compactenum}

\subsubsection*{Significant infrastructure}

{Both CASA and the ICMS have significant in-house computing resources (several computing clusters) as well as access to national computing facilities at SARA (Amsterdam).}
\end{sitedescription}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:
